%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[0]{Acknowledgments}{acknowledgments}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}

First and fore most I would like to praise God, the almighty, for His continuous provision of grace and unfailing mercy through out my study and life overall. \\

Secondly, I would like to express my sincere gratitude towards my thesis supervisors Dr. Ing. Frank Krüger and David Schindler for offering me an opportunity to work under their supervision. I appreciate every precious time, effort and resources they have invested in me. They have been extremely  helpful through out my thesis work and I could not imagine doing this project without their immense support and direction. \\

I would also like to thank my parents, Tadesse Kobro and Debritu Berhanu, for their continuous support and sacrifices. My gratitude further goes to my siblings: Bamlak, Japheth, Hannah and Jerusalem for their understanding and love.   I also feel a deep sense of gratitude towards my uncle Netsanet Berhanu for being a continuous source of wisdom and inspiration. \\

Last but not least, I owe a heartly gratitude to my pastor and friend, Kevan Franklin Smith and his wife Pricilla for their empathy, continuous prayers and support.  I would love to appreciate pastor Kevan for his profound wisdom and counseling. 



\bigskip


\endgroup



